---
title: "Barcode preprocessing - Prime reporter library"
author: 
  - name: "Max Trauernicht"
    email: "m.trauernicht@nki.nl"
    affiliation: "Netherlands Cancer Institute - van Steensel lab"
date: '`r format(Sys.time(), "%d/%m/%Y")`'
output: 
  html_document:
    theme: united
    highlight: pygments
    fig_caption: yes
    code_folding: hide
    df_print: kable
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
---
---

### Aim
In this script I want to load Sanger sequencing files from colony PCRs and match those to the list of prime reporters to identify individual prime TF reporters.

---

## Setup {.tabset}

<!-- little HTML script to do indentation of the table of contents -->
<script>
    $(document).ready(function() {
      $items = $('div#TOC li');
      $items.each(function(idx) {
        num_ul = $(this).parentsUntil('#TOC').length;
        $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
      });

    });
</script>

```{css, echo = FALSE}
div.sourceCode {
  overflow-x: hidden;
}
```

### Libraries 

```{r setup, out.width= "80%", fig.align= "center",  warning= FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(plotly)
library(ggpubr)
library(plyr)
library(data.table)
library(ggbeeswarm)
library(viridis)
library(GGally)
library(pheatmap)
library(sangerseqR)
```


### Functions

```{r out.width= "80%", fig.align= "center",  warning= FALSE, message=FALSE}
# Custom functions
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}

# From Fede:
# ggpairs custom functions
corColor <- function(data, mapping, color = I("black"), sizeRange = c(1, 3), ...) {

  x   <- eval_data_col(data, mapping$x)
  y   <- eval_data_col(data, mapping$y)
  r   <- cor(x, y, "pairwise.complete.obs")
  rt  <- format(r, digits = 3)
  tt  <- as.character(rt)
  cex <- max(sizeRange)

  # helper function to calculate a useable size
  percent_of_range <- function(percent, range) {
    percent * diff(range) + min(range, na.rm = TRUE)
  }

  # plot correlation coefficient
  p <- ggally_text(label = tt, mapping = aes(), xP = 0.5, yP = 0.5,
                   size = I(percent_of_range(cex * abs(r), sizeRange)), color = color, ...) +
    theme(panel.grid.minor=element_blank(),
          panel.grid.major=element_blank())

  corColors <- RColorBrewer::brewer.pal(n = 7, name = "RdYlBu")[2:6]

  if (r <= boundaries[1]) {
    corCol <- corColors[1]
  } else if (r <= boundaries[2]) {
    corCol <- corColors[2]
  } else if (r < boundaries[3]) {
    corCol <- corColors[3]
  } else if (r < boundaries[4]) {
    corCol <- corColors[4]
  } else {
    corCol <- corColors[5]
  }

  p <- p +
    theme(panel.background = element_rect(fill = corCol))

  return(p)
}

# Custom ggplot2 themes
theme_classic_lines <- function() {
  theme_pubr(border = T, legend = "top") +
            theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.1),
                  strip.background = element_rect(fill = "#ced4da"))
    
}

theme_classic_lines_45 <- function() {
  theme_pubr(border = T, legend = "top", x.text.angle = 45) +
            theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.1),
                  strip.background = element_rect(fill = "#ced4da"))
}

theme_classic_lines_90 <- function() {
  theme_pubr(border = T,legend = "top", x.text.angle = 90) +
            theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.1),
                  strip.background = element_rect(fill = "#ced4da"))
}

theme_set(theme_classic_lines())

colors_diverse <- c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51")

ggplot_custom <- function(...) ggplot2::ggplot(...) + 
  scale_color_manual(values = colors_diverse) + 
  scale_fill_manual(values = colors_diverse)
```

### Loading data

```{r data import, out.width= "80%", fig.align= "center",  warning= FALSE, message=FALSE}
# Load prime reporter library
prime_reporters <- read.csv("/DATA/usr/m.trauernicht/projects/prime_TF_reporters/library_design/prime_reporter_library.csv", header = TRUE, sep = ",") %>%
  dplyr::select(sequence, "reporter_id" = tf_bc) %>%
  mutate(barcode = gsub("CATCGTCGCATCCAAGAG", "", sequence)) %>%
  mutate(barcode = gsub(".*([A-Z]{13})", "\\1", barcode))
```


```{r data import, out.width= "80%", fig.align= "center",  warning= FALSE, message=FALSE}
# Load sequences from .ab1 files
folder_path <- "/DATA/usr/m.trauernicht/projects/prime_TF_reporters/data/sanger"
ab1_files <- list.files(folder_path, pattern = "\\.ab1$", full.names = TRUE)

# Function to read an ab1 file and extract the primary sequence
read_ab1_sequence <- function(file) {
  tryCatch({
    ab1_data <- readsangerseq(file)
    return(primarySeq(ab1_data))
  }, error = function(e) {
    message(paste("Error reading file:", file))
    return(NULL)  # Return NULL if there's an error
  })
}
# Apply the function to all ab1 files
sequences <- lapply(ab1_files, read_ab1_sequence)

# Create a data frame but first filter out invalid sequences
sequences_df <- data.frame(
  file_name = basename(ab1_files),  # Add file names as a column
  sequence = sapply(sequences, function(seq) {
    if(length(as.character(seq)) > 0) {
      as.character(seq)  # Extract sequence if it's non-empty
    } else {
      NA  # Mark invalid sequences as NA
    }
  }),
  stringsAsFactors = FALSE
)

sequences_df_filt <- sequences_df %>%
  filter(sequence != "NNNNN", str_detect(file_name, "Gibson", negate = T)) %>%
  mutate(barcode = gsub("CATCGTCGCATCCAAGAG.*", "", sequence)) %>%
  mutate(barcode = gsub(".*([A-Z]{13})", "\\1", barcode))
```

### Match sequences to reporter library

```{r}
sequences_df_filt_match <- sequences_df_filt %>%
  left_join(prime_reporters %>% dplyr::select(-sequence), by = "barcode")
```

