---
title: "Detailed TF reporter activity analysis"
author: 
  - name: "Max Trauernicht"
    email: "m.trauernicht@nki.nl"
    affiliation: "Netherlands Cancer Institute - van Steensel lab"

date: '`r format(Sys.time(), "%d/%m/%Y")`'
output:
  html_document:
    theme: united
    highlight: pygments
    fig_caption: yes
    code_folding: hide
    df_print: kable
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
---

---

### Aim
35,500 reporters for 86 TFs were transfected into 9 different cell types + tested across ~100 perturbation conditions. In this script I will analyze TF reporter activities in detail and review how individual reporters respond to TF perturbations.


### Description of input data
- commercial_reporter = positive control reporter sequence from commercial source
- neg_ctrls = reporter with mutated TF binding sites?
- hPGK = positive control, chunk from hPGK promoter (only in lib 1)
- native_enhancer = genomic mES enhancer sequence chunks from Miguels libraries (only lib 1)
- reporter_activity_minP = mean activity relative to median of negative control reporters (per core promoter)
- reporter_activity_neg = mean activity relative to its paired negative control (identical reporter with mutated binding sites)

---

## Setup {.tabset}

<!-- little HTML script to do indentation of the table of contents -->
<script>
    $(document).ready(function() {
      $items = $('div#TOC li');
      $items.each(function(idx) {
        num_ul = $(this).parentsUntil('#TOC').length;
        $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
      });

    });
</script>

```{css, echo = FALSE}
div.sourceCode {
  overflow-x: hidden;
}
```

### Libraries 

```{r setup, out.width= "80%", fig.align= "center", warning = FALSE, message= FALSE}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(RColorBrewer)
library(ggplot2)
library(dplyr)
library(maditr)
library(tibble)
library(pheatmap)
library(ggpubr)
library(visNetwork)
library(ggbeeswarm)
library(ggforce)
library(viridis)
library(plyr)
library(igraph)
library(ggraph)
library(cowplot)
library(gridExtra)
library(pROC)
library(tidyr)
library(stringr)
library(randomForest)
library(ggrastr)
library(readr)
library(ggbiplot)
library(IHW)
library(biomaRt)
library(ggh4x)
library(ggiraph)
library(plotly)
library(AMR)
library(umap)
library(ggrepel)
```

### Functions

```{r out.width= "80%", fig.align= "center"}
### Custom functions
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}


# Extract p-value from linear model
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}

# Set custom ggplot2 theme and custom colors
theme_classic_lines <- function() {
  theme_pubr(border = F, legend = "top") +
            theme(panel.grid.major = element_line(colour = "#adb5bd", linewidth = 0.25),
                  strip.background = element_rect(fill = "#ced4da")
            )
}

theme_classic_lines_45 <- function() {
  theme_pubr(border = T, legend = "top", x.text.angle = 45) +
            theme(panel.grid.major = element_line(colour = "#adb5bd", linewidth = 0.25),
                  strip.background = element_rect(fill = "#ced4da")
            )
}

theme_classic_lines_90 <- function() {
  theme_pubr(border = T, legend = "top", x.text.angle = 90) +
            theme(panel.grid.major = element_line(colour = "#adb5bd", linewidth = 0.25),
                  strip.background = element_rect(fill = "#ced4da")
            )
}

theme_barplot <- function() {
  theme_pubr(border = T, legend = "none") +
            theme(panel.grid.major.x = element_line(colour = "black", linewidth = 0.25),
                  strip.background = element_blank(),
                  strip.text = element_text(face="bold", hjust=0)
            )
}

theme_set(theme_classic_lines())

colors_diverse <- c("#264653", "#9AC1AE", "#5D987B", "#f2cc8f", "#e76f51")

cell_colors <- c("A549" = "#f2cc8f", "HCT116" = "#ED1C24", "HEK293" = "#00B4D8", 
                               "HepG2" = "#B3B3B3", "K562" = "#A67C52", "MCF7" = "#81B29A", 
                               "U2OS" = "#3D405B", "mES" = "#EAB69F", "NPC" = "#E07A5F")

ggplot_custom <- function(...) ggplot2::ggplot(...) + 
  scale_color_manual(values = colors_diverse) + 
  scale_fill_manual(values = colors_diverse)
```

### Loading data

```{r out.width= "80%", fig.align= "center"}
# Import processed bc counts from the preprocessing step
cDNA_df <- read.csv("/DATA/usr/m.trauernicht/projects/prime_TF_reporters/data/dan_20241003/mt20241004_reporter_activities.csv", header = T)

# Compute perturbation-induced fold-change
ref_activities <- cDNA_df %>%
  filter(condition == "PC9_DMSO") %>%
  dplyr::select("ref_activity" = mean_activity_sample_minP, replicate, reporter_id) %>%
  distinct()
  
cDNA_df <- cDNA_df %>%
  left_join(ref_activities, by = c("replicate", "reporter_id")) %>%
  mutate(mean_activity_dif = mean_activity_sample_minP / ref_activity) %>%
  mutate(reporter_activity_dif = ave(mean_activity_dif, condition, reporter_id, FUN = function(x) mean(x, na.rm = T))) %>%
  mutate(tf_activity_dif = ave(mean_activity_dif, condition, tf, FUN = function(x) mean(x, na.rm = T))) %>%
  mutate(tf = gsub("_.*", "", tf))

```

---

```{r out.width= "80%", fig.align= "center"}
## Plot the reporter activities of one TF of interest (replicates as individual dots)
ggplot(cDNA_df %>%
         filter(neg_ctrls == "No", native_enhancer == "No") %>%
         filter(tf %in% c("ELK1")) %>%
         mutate(mean_activity_sample_minP = log2(mean_activity_sample_minP)) %>%
         distinct(mean_activity_sample_minP, reporter_id, condition, tf, replicate, prime),
       aes(x = condition, y = mean_activity_sample_minP, color = prime)) +
  geom_quasirandom() +
  scale_color_manual(values = c("yes" = "red", "no" = "black")) +
  ylab("Activity (log2)") +
  theme_pubr(x.text.angle = 90)

## Plot the perturbation-induced fold-change of one TF of interest (replicates separately)
ggplot(cDNA_df %>%
         filter(neg_ctrls == "No", native_enhancer == "No", !is.na(stimulation)) %>%
         filter(tf %in% c("ELK1")) %>%
         distinct(mean_activity_dif, reporter_id, condition, tf, replicate, prime),
       aes(x = replicate, y = mean_activity_dif, color = prime)) +
  geom_hline(yintercept = 1, lty = 2) +
  geom_quasirandom() +
  scale_color_manual(values = c("yes" = "red", "no" = "black")) +
  ylab("Activity (log2)") +
  theme_pubr(x.text.angle = 90)

## Plot the perturbation-induced fold-change of all TFs (with a fold-change > 1.5)
tfs_changed <- cDNA_df %>%
  filter(tf_activity_dif > 1.5 | tf_activity_dif < 1/1.5)

ggplot(cDNA_df %>%
         filter(tf %in% tfs_changed$tf, !is.na(stimulation)) %>%
         filter(neg_ctrls == "No", native_enhancer == "No") %>%
         distinct(mean_activity_dif, reporter_id, tf, prime, replicate, commercial_reporter),
       aes(x = replicate, y = mean_activity_dif, color = prime, shape = commercial_reporter)) +
  geom_hline(yintercept = 1, lty = 2) +
  geom_quasirandom() +
  scale_color_manual(values = c("yes" = "red", "no" = "black")) +
  ylab("Activity (log2)") +
  theme_pubr(x.text.angle = 90, border = T) +
  facet_wrap(~tf, scales = "free_y")
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

