---
title: "Oligo Design - TF reporter screen gen 2"
author: "Max Trauernicht"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  prettydoc::html_pretty:
    theme: leonids
  #   highlight: github
  #   toc: true
  #   toc_float: true
  #   code_folding: show
  # editor_options:
  #   chunk_output_type: console
---

*knitr document van Steensel lab*

# Oligo Design - TF reporter screen gen 2

## Introduction
In this document, an oligo library of TF reporters will be generated. These oligos contain TF binding sites of ~30 selected TFs, random inactive spacing between them, followed by minimal promoters, and a barcode in the transcription unit.

### Load libraries

```{r setup, out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(ggplot2)
library(seqinr)
library(seqLogo)
library(universalmotif)
library(Biostrings)
library(SimRAD)
library(gtools)
library(DNABarcodes)
library(phylotools)
library(ape)
library(magrittr)
library(dplyr)
library(readr)
library(stringr)
library(tidyr)
library(heatmaply)
library(pheatmap)
library(tibble)
library(ggseqlogo)
library(RColorBrewer)
library(data.table)
library(vwr)
library(ggbeeswarm)
library(ggrepel)
library(sjmisc)
library(RcmdrMisc)
library(LncFinder)
library(ENCODExplorer)
library(biomaRt)
library(ggpubr)
```

### Load functions

```{r out.width= "100%", fig.align= "center", echo=FALSE, warning= FALSE, message=FALSE}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}

# Function to substring the right part of the motif
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}


# Function to load PWM matrix
get_pwm_feature_matrix <- function(motif_meta_fn, fimo_fn, db = 2) {

  # validate args
  valid_dbs <- 1:2
  if(!db %in% valid_dbs)
    stop('Invalid db (database version). Please use db=1 (maintained for backward compatibility only) or db=2')

  # db=1 is maintained for backward compatibility only
  if(db == 1) {

    # read in motif metadata
    motif_meta    <- read.csv(motif_meta_fn)

    # check whether motif metadata contain essential annotations
    if(!all(c('PWM.ID', 'Cognate.TF') %in% colnames(motif_meta))) {
      message('The motif metadata file does not contain the essential columns PWM.ID and Cognate.TF')
    }

    motif_minimal <- motif_meta[, c('PWM.ID', 'Cognate.TF')]

    # load fimo output --> extract motif id, sequence id and p-value
    df <- read.table(fimo_fn)
    df <- df[, c(1, 2, 7)]

    colnames(df) <- c('PWM.ID', 'seqid', 'pval')

    # add TF id
    df <- merge(df, motif_minimal, by = 'PWM.ID')

    # group motif hits by sequence id
    l <- split(df, df[['seqid']])

    # multiple PWM and multiple hits possible. Reduce hits to one per TF, keeping best p-val only
    l <- lapply(l, function(x) {
      x_by_tf <- split(x, x[['Cognate.TF']], drop = TRUE)
      x_by_tf <- lapply(x_by_tf, function(y) y[which.min(y$pval), ])
      do.call('rbind', x_by_tf)
    })

    # initialize feature matrix
    n_tf          <- motif_minimal[['Cognate.TF']] %>%
      unique %>%
      length
    n_seq         <- length(l)
    pwm           <- matrix(1, nrow = n_seq, ncol = n_tf)
    colnames(pwm) <- (motif_minimal[['Cognate.TF']] %>% unique)

    # replace :: from names of composite motifs
    colnames(pwm) <- str_replace_all(colnames(pwm), '::', '_')

    # fill in feature matrix
    for(i in 1 : n_seq) {
      pwm[i, l[[i]][['Cognate.TF']]] <- l[[i]]$pval
    }

    # -log10 transform
    pwm           <- -1 * log10(pwm)

    # coerce to tib and return
    tib_fimo <- as_data_frame(pwm) %>%
      mutate(id = names(l))
      dplyr::select(id, everything())

  }

  # db = 2 (default)
  else {

    # load metadata
    tib_meta    <- read_csv(motif_meta_fn) %>%
      # extract tf symbol from motif id (Cognate_TF unsafe, it can be empty) and replace :: occurrences
      mutate(tf_symbol = str_remove(ID, '_[0-9]*'),
             tf_symbol = str_replace(tf_symbol, '::', '_')) %>%
      dplyr::select(motif_id = `PWM ID`, tf_symbol)

    # load fimo results
    tib_fimo <- read_tsv(fimo_fn) %>%
      # extract motif id, sequence id and p-value
      dplyr::select(motif_id, sequence_name, pval = `p-value`, start)

    # add tf symbol to fimo results
    tib_fimo <- tib_fimo %>%
      left_join(tib_meta, by = 'motif_id') %>%
      # remove hits with missing motif id (composite pwms)
      filter(!is.na(tf_symbol))

    # # select best hit for each motif and sequence
    # tib_fimo <- tib_fimo %>%
    #   group_by(sequence_name, tf_symbol) %>%
    #   dplyr::slice(which.min(pval)) %>%
    #   ungroup()

    # spread into feature matrix
    tib_fimo <- tib_fimo %>%
      mutate(pval = -1 * log10(pval)) %>%
      dplyr::select(-motif_id) %>%
      spread(key = tf_symbol, value = pval, fill = 0, drop = TRUE) %>%
      # perform cosmetics on the id
      mutate(id = sequence_name) %>%
      dplyr::select(-c(sequence_name)) %>%
      dplyr::select(id, everything())

  }

  return(tib_fimo)

}

ReadFasta<-function(file) {
   # Read the file line by line
   fasta<-readLines(file)
   # Identify header lines
   ind<-grep(">", fasta)
   # Identify the sequence lines
   s<-data.frame(ind=ind, from=ind+1, to=c((ind-1)[-1], length(fasta)))
   # Process sequence lines
   seqs<-rep(NA, length(ind))
   for(i in 1:length(ind)) {
      seqs[i]<-paste(fasta[s$from[i]:s$to[i]], collapse="")
   }
   # Create a data frame 
   DF<-data.frame(name=gsub(">", "", fasta[ind]), sequence=seqs)
   # Return the data frame as a result object from the function
   return(DF)
}
```


### Import data
```{r}
# Load from TF_reporter_variant_analysis.Rmd
best_reporters <- read_csv("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/best_reporters.csv")

# Load sequences from library design files
lib1_seqs <- read.fasta("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/library_design/output/mt20191205_oligo_pool.fasta")
lib2_seqs <- read.fasta("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/library_design/gen-2/mt20210326_oligo_pool_gen2.fasta")

# Combine libraries and extract barcodes to match with other file
lib1_seqs <- lib1_seqs %>%
  mutate(barcode = gsub(".*CACGACGCTCTTCCGATCT([A-Z]{12}).*", "\\1", seq.text))

lib2_seqs <- lib2_seqs %>%
  mutate(barcode = gsub(".*CACGACGCTCTTCCGATCT([A-Z]{13}).*", "\\1", seq.text))

lib_1_2_seqs <- lib1_seqs %>%
  rbind(lib2_seqs) %>%
  dplyr::select("sequence" = seq.text, barcode)

# Import barcode annotation of both libraries
bc_annotation <- read_csv("/DATA/usr/m.trauernicht/projects/SuRE-TF/data/library_design/bc_annotation_combined.csv") %>%
  filter(library == "1+2") %>%
  mutate(reporter_id = paste(tf, spacing, distance, promoter, background, sep = "_"))

# Combine sequences and best reporter IDs
reporter_seqs <- bc_annotation %>%
  distinct(barcode, reporter_id, barcode_number) %>%
  left_join(lib_1_2_seqs) %>%
  distinct(reporter_id, sequence, barcode_number)

best_reporters_seq <- best_reporters %>%
  distinct(tf, reporter_id, conf_level) %>%
  left_join(reporter_seqs) %>%
  filter(barcode_number <= 5)

## Manually reduce the length of the barcodes of NR1H2 and STAT6 reporters (so the total length of the library is max. 250bp)
best_reporters_seq_nr1h2_stat6 <- best_reporters_seq %>%
  filter(tf %in% c("NR1H2", "STAT6")) %>%
  mutate(left = gsub("(.*CACGACGCTCTTCCGATCT).*", "\\1", sequence)) %>%
  mutate(right = gsub(".*CACGACGCTCTTCCGATCT(.*)", "\\1", sequence)) %>%
  mutate(right = ifelse(tf == "NR1H2", substr(right, 2, 31), right)) %>%
  mutate(right = ifelse(tf == "STAT6", substr(right, 5, 31), right)) %>%
  mutate(sequence = paste(left, right, sep = "")) %>%
  dplyr::select(-c(left, right))

## Manually include a fifth NFAT5 reporter (somehow one was missing - I will just take a barcode from another non-prime NFAT5 reporter)
best_reporters_seq_nfat5 <- best_reporters_seq %>%
  filter(tf %in% c("NFAT5"), barcode_number == 2) %>%
  mutate(barcode_number = 1) %>%
  mutate(sequence = gsub("(.*CACGACGCTCTTCCGATCT)[A-Z]{13}(.*)", "\\1GTCACCAAGTTGG\\2", sequence))

best_reporters_seq <- best_reporters_seq %>%
  filter(tf != "NR1H2", tf != "STAT6") %>%
  rbind(best_reporters_seq_nr1h2_stat6) %>%
  rbind(best_reporters_seq_nfat5)

# Include random reporters -  all three scrambled TFBSs - only one design per core promoter
bc_annotation_random <- bc_annotation %>%
  filter(reporter_id %in% c("RANDOM1_10bp_21bp_minP_1", "RANDOM1_10bp_21bp_mCMV_1", "RANDOM1_10bp_21bp_hBGm_1", "RANDOM1_10bp_21bp_Random_1",
                            "RANDOM1_10bp_21bp_minP_2", "RANDOM1_10bp_21bp_mCMV_2", "RANDOM1_10bp_21bp_hBGm_2", 
                            "RANDOM1_10bp_21bp_minP_3", "RANDOM1_10bp_21bp_mCMV_3", "RANDOM1_10bp_21bp_hBGm_3", 
                            "RANDOM2_10bp_21bp_minP_1", "RANDOM2_10bp_21bp_mCMV_1", "RANDOM2_10bp_21bp_hBGm_1", "RANDOM2_10bp_21bp_Random_1",
                            "RANDOM2_10bp_21bp_minP_2", "RANDOM2_10bp_21bp_mCMV_2", "RANDOM2_10bp_21bp_hBGm_2", 
                            "RANDOM2_10bp_21bp_minP_3", "RANDOM2_10bp_21bp_mCMV_3", "RANDOM2_10bp_21bp_hBGm_3", 
                            "RANDOM3_10bp_21bp_minP_1", "RANDOM3_10bp_21bp_mCMV_1", "RANDOM3_10bp_21bp_hBGm_1", "RANDOM3_10bp_21bp_Random_1",
                            "RANDOM3_10bp_21bp_minP_2", "RANDOM3_10bp_21bp_mCMV_2", "RANDOM3_10bp_21bp_hBGm_2", 
                            "RANDOM3_10bp_21bp_minP_3", "RANDOM3_10bp_21bp_mCMV_3", "RANDOM3_10bp_21bp_hBGm_3")) %>%
  filter(barcode_number <= 5)

reporter_seqs_random <- bc_annotation_random %>%
  distinct(barcode, reporter_id, barcode_number, tf) %>%
  left_join(lib_1_2_seqs) %>%
  distinct(reporter_id, sequence, barcode_number, tf) %>%
  # Add a column that counts the rows per reporter_id
  dplyr::group_by(reporter_id) %>%
  dplyr::mutate(barcode_number2 = dplyr::row_number()) %>%
  dplyr::ungroup() %>%
  filter(barcode_number2 <= 5) %>%
  dplyr::select(-barcode_number2)


# Extract sequences
best_reporters_seq2 <- best_reporters_seq %>%
  dplyr::select(-conf_level) %>%
  rbind(reporter_seqs_random) %>%
  mutate(tf_bc = paste(reporter_id, barcode_number, sep = "_"))

## Add the same reporters with only one barcode (the barcode with the activity closest to the median)  

### Import barcodes closest to median from barcode preprocessing script
median_barcode <- read_csv("/DATA/usr/m.trauernicht/projects/SuRE-TF/library_design/prime_reporter_library/median_barcodes.csv")

best_reporters_seq2_single <- best_reporters_seq2 %>%
  right_join(median_barcode) %>%
  mutate(sequence = gsub("", "", sequence)) %>%
  mutate(sequence = gsub("", "", sequence)) %>%
  mutate(tf_bc = paste(tf_bc, "single", sep = "_"))

best_reporters_seq_export <- best_reporters_seq2 %>%
  distinct(tf_bc, sequence)

best_reporters_seq2_single_export <- best_reporters_seq2_single %>%
  distinct(tf_bc, sequence)

write_csv(best_reporters_seq_export, "/DATA/usr/m.trauernicht/projects/SuRE-TF/library_design/prime_reporter_library/prime_reporter_library.csv")
write_csv(best_reporters_seq2_single_export, "/DATA/usr/m.trauernicht/projects/SuRE-TF/library_design/prime_reporter_library/prime_reporter_library_single.csv")
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

